---
title: "cross_validation"
author: "ruby"
date: "2025-11-11"
output: html_document
---

load packages and dataset 

```{r message = FALSE}
library(tidyverse)
library(modelr)
library(mgcv)

library(p8105.datasets)
set.seed(1)



knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

cross-validation - ML concept, testing and training models 

load LIDAR 

```{r}
data("lidar")
```

look at the `lidar` dataset

```{r}
lidar

lidar_df = 
  lidar |> 
  mutate(id = row_number())


lidar_df |> 
  ggplot(aes(x = range, y = logratio)) +
  geom_point()
```

create dataframes 

```{r}
train_df = 
  sample_frac(lidar_df, size = .8) |> 
  arrange(id)

test_df = 
  anti_join(lidar_df, train_df, by = "id")
```

taking 80% of dataset in training dataframe 
testing dataframe has the rows that are in the lidar_df that are not in the train_df

* essentially taking the full dataset and splitting it into 2 df (testing and training)

look at these 
```{r}
ggplot(train_df, aes(x = range, y = logratio)) +
  geom_point() +
  geom_point(data = test_df, color = "red")


```

fit a few models to `train_df`


```{r}
linear_mod = 
  lm(logratio ~ range, data = train_df)

smooth_mod = 
  mgcv::gam(logratio ~ s(range), data = train_df)

wiggly_mod = 
  mgcv::gam(logratio ~ s(range, k = 50), sp = 10e-8, data = train_df)
```

look at this 

```{r}
train_df |> 
  add_predictions(linear_mod) |> 
  ggplot(aes(x = range, y = logratio)) +
  geom_point() +
  geom_line(aes(y = pred), color = "red")


train_df |> 
  add_predictions(smooth_mod) |> 
  ggplot(aes(x = range, y = logratio)) +
  geom_point() +
  geom_line(aes(y = pred), color = "red")


train_df |> 
  add_predictions(wiggly_mod) |> 
  ggplot(aes(x = range, y = logratio)) +
  geom_point() +
  geom_line(aes(y = pred), color = "red")
```

you should not make a `wiggly` model! this will actually make slightly less accurate predictions than a smooth model - overfitting model


