---
title: "cross_validation"
author: "ruby"
date: "2025-11-11"
output: html_document
---

load packages and dataset 

```{r message = FALSE}
library(tidyverse)
library(modelr)
library(mgcv)

library(p8105.datasets)
set.seed(1)



knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

cross-validation - ML concept, testing and training models 

load LIDAR 

```{r}
data("lidar")
```

look at the `lidar` dataset

```{r}
lidar

lidar_df = 
  lidar |> 
  mutate(id = row_number())


lidar_df |> 
  ggplot(aes(x = range, y = logratio)) +
  geom_point()
```

create dataframes 

```{r}
train_df = 
  sample_frac(lidar_df, size = .8) |> 
  arrange(id)

test_df = 
  anti_join(lidar_df, train_df, by = "id")
```

taking 80% of dataset in training dataframe 
testing dataframe has the rows that are in the lidar_df that are not in the train_df

* essentially taking the full dataset and splitting it into 2 df (testing and training)

look at these 
```{r}
ggplot(train_df, aes(x = range, y = logratio)) +
  geom_point() +
  geom_point(data = test_df, color = "red")


```

fit a few models to `train_df`


```{r}
linear_mod = 
  lm(logratio ~ range, data = train_df)

smooth_mod = 
  mgcv::gam(logratio ~ s(range), data = train_df)

wiggly_mod = 
  mgcv::gam(logratio ~ s(range, k = 50), sp = 10e-8, data = train_df)
```

look at this 

```{r}
train_df |> 
  add_predictions(linear_mod) |> 
  ggplot(aes(x = range, y = logratio)) +
  geom_point() +
  geom_line(aes(y = pred), color = "red")


train_df |> 
  add_predictions(smooth_mod) |> 
  ggplot(aes(x = range, y = logratio)) +
  geom_point() +
  geom_line(aes(y = pred), color = "red")


train_df |> 
  add_predictions(wiggly_mod) |> 
  ggplot(aes(x = range, y = logratio)) +
  geom_point() +
  geom_line(aes(y = pred), color = "red")
```

you should not make a `wiggly` model! this will actually make slightly less accurate predictions than a smooth model - overfitting model

try computing our RMSEs 

```{r}
rmse(linear_mod, test_df)
rmse(smooth_mod, test_df)
rmse(wiggly_mod, test_df)
```

## iterate!

```{r}
cv_df = 
  crossv_mc(lidar_df, n = 100) |> 
  mutate(
    train = map(train, as_tibble),
    test = map(test, as_tibble)
  )
```

did this work? (did the 80/20 split) `crossv_mc` uses an 80/20 fit by default

```{r}
cv_df |>  
  pull(train) |> 
  nth(1)
```

fit models over and over. (`map2` maps over 2 things, linear fits and test)

```{r}
cv_df = 
  cv_df |> 
    mutate(
      linear_fits = map(train, \(df) lm(logratio ~ range, data = df)),
      smooth_fits = map(train, \(df) mgcv::gam(logratio ~ s(range), data = df)),
      wiggly_fits = map(train, \(df) mgcv::gam(logratio ~ s(range, k = 50), sp = 10e-8, data = df))
    ) |> 
    mutate(
      rmse_linear = map2_dbl(linear_fits, test, rmse),
      rmse_smooth = map2_dbl(smooth_fits, test, rmse),
      rmse_wiggly = map2_dbl(wiggly_fits, test, rmse)
    )

```

thw anonymous function above is the same as writing the function separately (below) and then using this map statement

`map(train, lidar_lm, data = df)`

```{r eval = FALSE }
lidar_lm = function(df) {
  
  lm(logratio ~ range, data = df)
}
```


let's try to look at this better 

```{r}
cv_df |> 
  select(starts_with("rmse")) |> 
  pivot_longer(
    everything(),
    names_to = "model", 
    values_to = "rmse",
    names_prefix = "rmse_"
  ) |> 
  ggplot(aes(x = model, y = rmse)) +
  geom_violin()
```

### child growth dataset example 

```{r}
growth_df = 
  read_csv("./data/nepalese_children.csv")
```

weight vs. armc

```{r}
growth_df |> 
  ggplot(aes(x = weight, y = armc)) +
  geom_point(alpha = .5)
```


make a cut off point at weight = 7
```{r}
growth_df = 
  growth_df |> 
  mutate(
    weight_cp7 = (weight > 7) * (weight - 7)
  )
```

fitting 3 models 

linear model, piecewise model, and smooth model
```{r}
linear_mod = lm(armc ~ weight, growth_df)

pwl_mod = lm(armc ~ weight + weight_cp7, growth_df)

smooth_mod = mgcv::gam(armc ~ s(weight), data= growth_df)
```

```{r}
growth_df |> 
  add_predictions(linear_mod) |> 
  ggplot(aes(x = weight, y = armc)) +
  geom_point(alpha = 0.5) +
  geom_line(aes(y = pred), color = "red")

growth_df |> 
  add_predictions(pwl_mod) |> 
  ggplot(aes(x = weight, y = armc)) +
  geom_point(alpha = 0.5) +
  geom_line(aes(y = pred), color = "red")

growth_df |> 
  add_predictions(smooth_mod) |> 
  ggplot(aes(x = weight, y = armc)) +
  geom_point(alpha = 0.5) +
  geom_line(aes(y = pred), color = "red")
```

cross-validate to see which of these models is the best at making predictions 

```{r}
cv_df = 
  crossv_mc(growth_df, n = 100) |> 
  mutate(
    train = map(train, as_tibble),
    test = map(test, as_tibble)
  )
```

```{r}
cv_df = 
  cv_df |> 
  mutate(
    linear_mod = map(train, \(df) lm(armc ~ weight, data = df)),
    pwl_mod = map(train, \(df) lm(armc ~ weight + weight_cp7, data = df)),
    smooth_mod = map(train, \(df) mgcv::gam(armc ~ s(weight), data = df))
  ) |> 
  mutate(
    rmse_linear = map2_dbl(linear_mod, test, rmse),
    rmse_pwl = map2_dbl(pwl_mod, test, rmse),
    rmse_smooth = map2_dbl(smooth_mod, test, rmse)
  )
```

create some plots 

```{r}
cv_df |> 
  select(starts_with("rmse")) |> 
  pivot_longer(
    everything(),
    names_to = "model", 
    values_to = "rmse",
    names_prefix = "rmse_"
  ) |> 
  ggplot(aes(x = model, y = rmse)) +
  geom_violin()
```

