---
title: "linear_models"
author: "ruby"
date: "2025-11-06"
output: github_document
---

```{r message = FALSE}
library(tidyverse)
library(p8105.datasets)

set.seed(1)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

```

load the nyc airbnb data 

```{r}
data("nyc_airbnb")
```

look at the data/do some cleaning. 

```{r}
nyc_airbnb = 
  nyc_airbnb |> 
  mutate(
    stars = review_scores_location / 2
  ) |> 
  rename(
    borough = neighbourhood_group
  ) |> 
  filter(borough != "Staten Island") |> 
  select(price, stars, borough, room_type, neighbourhood)
```

do regression

```{r}
fit = lm(price ~ stars + borough, data = nyc_airbnb)
```

do some additional cleaning then refit. (how does changing which borough is the reference group change the linear model)

```{r}
nyc_airbnb = 
  nyc_airbnb |> 
  mutate(
    borough = fct_infreq(borough),
    room_type = fct_infreq(room_type))

fit = lm(price ~ stars + borough, data = nyc_airbnb)
```

look at `lm` stuff 

```{r, eval = FALSE} 
summary(fit)
names(summary(fit))
summary(fit)[["coefficients"]]
summary(fit)[["df"]]

fitted.values(fit)
```

look at cleaner `lm` stuff

```{r}
fit |> 
  broom::tidy() |> 
  mutate(
    term = str_replace(term, "borough", "Borough: ")
  ) |> 
  select(term, estimate, p.value) |> 
  knitr::kable(digits = 3)


fit |> 
  broom::glance()
```


## Diagnostics 

look at residuals.

```{r}
nyc_airbnb |> 
  modelr::add_residuals(fit) |> 
  modelr::add_predictions(fit) |> 
  filter(resid < 1000) |> 
  ggplot(aes(x = resid)) +
  geom_histogram()



nyc_airbnb |> 
  modelr::add_residuals(fit) |> 
  modelr::add_predictions(fit) |> 
  filter(resid < 1000) |> 
  ggplot(aes(x = borough, y = resid)) +
  geom_violin()


nyc_airbnb |> 
  modelr::add_residuals(fit) |> 
  modelr::add_predictions(fit) |> 
  filter(resid < 1000) |> 
  ggplot(aes(x = stars, y = resid)) +
  geom_point()
```

## Hypothesis testing 

```{r}
fit |> 
  broom::tidy()
```

What about a categorical variable?

```{r}
fit_alt = lm(price ~ stars + borough + room_type, data = nyc_airbnb)
fit_null = lm(price ~ stars + borough, data = nyc_airbnb)

anova(fit_null, fit_alt) |> 
  broom::tidy()
```

## Interactions vs nested data. 

```{r}
fit_interactions = 
  lm(price ~ stars * borough + room_type * borough, data = nyc_airbnb)

fit_interactions |> 
  broom::tidy()
```

let's just look at Brooklyn first. 

```{r}
nyc_airbnb |> 
  filter(borough == "Brooklyn") |> 
  lm(price ~ stars + room_type, data = _) |> 
  broom::tidy()

nyc_airbnb |> 
  filter(borough == "Manhattan") |> 
  lm(price ~ stars + room_type, data = _) |> 
  broom::tidy()

nyc_airbnb |> 
  filter(borough == "Bronx") |> 
  lm(price ~ stars + room_type, data = _) |> 
  broom::tidy()

nyc_airbnb |> 
  filter(borough == "Queens") |> 
  lm(price ~ stars + room_type, data = _) |> 
  broom::tidy()
```

Write a short function 

```{r}
lm_airbnb = function(df) {
  
  lm(price ~ stars + room_type, data = df)
  
}
  

nyc_airbnb |> 
  filter(borough == "Queens") |> 
  lm_airbnb() |> 
  broom::tidy()
```

Create a list of dataframes, and iterate ro fit the model each time. 

```{r}
nyc_airbnb |> 
  nest(data = -borough) |> 
  mutate(
    fits = map(data, lm_airbnb), 
    results = map(fits, broom::tidy)
  ) |> 
  select(borough, results) |> 
  unnest(results)
```

